---
title: "Deliverable 2"
author: "Júlia Gasull i Claudia Sánchez"
date: \today
output:
  html_document:
    toc: no
    toc_depth: '4'
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
  word_document:
    toc: no
    toc_depth: '4'
geometry: left=1.9cm,right=1.9cm,top=1.25cm,bottom=1.52cm
fontsize: 18pt
subtitle: PCA, CA and Clustering
classoption: a4paper
editor_options: 
  chunk_output_type: console
---

# First setups
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = T, results = 'hide'}
if(!is.null(dev.list())) dev.off()  # Clear plots
rm(list=ls())                       # Clean workspace
```

## Load Required Packages for this deliverable
We load the necessary packages and set working directory
```{r echo = T, results = 'hide', message=FALSE, error=FALSE, warning=FALSE}

setwd("~/Documents/uni/FIB-ADEI-LAB/deliverable2")
filepath<-"~/Documents/uni/FIB-ADEI-LAB/deliverable2"
#setwd("C:/Users/Claudia Sánchez/Desktop/FIB/TARDOR 2020-2021/ADEI/DELIVERABLE1/FIB-ADEI-LAB/deliverable2")
#filepath<-"C:/Users/Claudia Sánchez/Desktop/FIB/TARDOR 2020-2021/ADEI/DELIVERABLE1/FIB-ADEI-LAB/deliverable2"

# Load Required Packages
options(contrasts=c("contr.treatment","contr.treatment"))
requiredPackages <- c("missMDA","chemometrics","mvoutlier","effects","FactoMineR","car", "factoextra","RColorBrewer","dplyr","ggmap","ggthemes","knitr")
missingPackages <- requiredPackages[!(requiredPackages %in% installed.packages()[,"Package"])]
if(length(missingPackages)) install.packages(missingPackages)
lapply(requiredPackages, require, character.only = TRUE)
```

## Load processed data from first deliverable
```{r}
load(paste0(filepath,"/Taxi5000_del1.RData"))
summary(df)
```


## Clean data
```{r}
# remove some columns
df$lpep_pickup_datetime <- NULL
df$Lpep_dropoff_datetime <- NULL
df$Store_and_fwd_flag <- NULL
df$Ehail_fee <- NULL
df$CashTips <- NULL

# imputation
library(missMDA)
long_lat<-names(df)[c(3:6)]
imp_long_lat<-imputePCA(df[,long_lat])
df[,long_lat]<-imp_long_lat$completeObs

summary(df)
```


--------------------------------------------------------------------------------

# PCA analysis
Ja hem vist el profiling a l'entrega anterior. Per tant, ara, procedim a veure les components principals.
Recordem, el nostre target quantitatiu és Total_amount.
```{r}
library(FactoMineR)
names(df)
pca_vars<-names(df)[c(3:6,7:10,12,15,18,20:22,1:2,16,19,30)]; pca_vars
res.pca<-PCA(df[,pca_vars],quanti.sup=c(1:4,10),quali.sup=c(15:19))
```

## Eigenvalues and dominant axes analysis. How many axes we have to interpret according to Kaiser and Elbow's rule?

## Individuals point of view: Are they any individuals "too contributive"? To better understand the axes meaning use the extreme individuals. Detection of multivariant outliers and influent data.
## Interpreting the axes: Variables point of view coordinates, quality of representation, contribution of the variables  
## Perform a PCA taking into account also supplementary variables the supplementary variables can be quantitative and/or categorical 

# K-Means Classification
## Description of clusters
# Hierarchical Clustering
## Description of clusters

# CA analysis for your data should contain your factor version of the numeric target (previous) in K= 7 (maximum 10) levels and 2 factors:

## Eigenvalues and dominant axes analysis. How many axes we have to consider
## Are there any row categories that can be combined/avoided to explain the discretization of the numeric target.

# MCA analysis for your data should contain:

## Eigenvalues and dominant axes analysis. How many axes we have to consider for next Hierarchical Classification stage?
## Individuals point of view: Are they any individuals "too contributive"? Are there any groups?
## Interpreting map of categories: average profile versus extreme profiles (rare categories)
## Interpreting the axes association to factor map.
## Perform a MCA taking into account also supplementary variables (use all numeric variables) quantitative and/or categorical. How supplementary variables enhance the axis interpretation?

# Hierarchical Clustering (from MCA)

## Description of clusters
## Parangons and class-specific individuals.
## Comparison of clusters obtained after K-Means (based on PCA) and/or Hierarchical Clustering (based on PCA) focusing on Duration target.
## Comparison of clusters obtained after K-Means (based on PCA) and/or Hierarchical Clustering (based on PCA) focusing on the binary target.
